---
title: "Bipartite monolayer network example"
navbar:
  title: "The infomapecology package website"
  left:
    - text: "Home"
      href: index.html
    - text: "Monolayer"
      #href: monolayer/index.html
      menu:
        - text: "Bipartite monolayer network example"
          href: bipartite.html
        - text: "Monolayer directed network with hierarchical structure"
          href: monolayer_heirarchical.html  
    - text: "Monolayer directed network with node attributes"
      href: monolayer_node_attributes.html
    - text: "Flow models"
      href: flows.html
    - text: "Hypothesis testing"
      href: hypothesis_test.html
    - text: "Temporal multilayer network with global relax rates"
      href: multilayer_relax.html
    - text: "Temporal multilayer network with interlayer edges"
      href: multilayer_interlayer.html
output:
  html_document: 
    fig_height: 6
    fig_width: 9
    theme: flatly
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      number_sections: true

---
```{r klippy, echo=FALSE, include=TRUE}
# Use this to add a copy-code buttong
klippy::klippy(position = c('top', 'right'), color = 'darkred')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data set
A weighted bipartite network describing a plant-flower visitor interaction web (25 plant species and 79 flower visitor species) in the vicinity of Bristol, U.K.  To distinguish between the two node sets we number the pollinator species from 1-79 and the plants from 80-104. Data can be obtained using `data(memmott1999)` using package `bipartite` in `R`.

### Input
A [link list](https://www.mapequation.org/infomap/#InputLinkList) with columns `from`, `to` and `weight`. Because this is a bipartite network the `from` column can only contain nodes 1-79 and the `to` column nodes 84-104.

### R Code
```{r Bipartite_1}
# Import data
library(infomapecology)
library(bipartite)
data(memmott1999)

network_object <- create_monolayer_network(memmott1999, bipartite = T, directed = F, group_names = c('A','P'))
infomap_object <- run_infomap_monolayer(network_object, infomap_executable='Infomap',
flow_model = 'undirected',
silent=T, trials=20, two_level=T, seed=123)

# Plot the matrix (plotting function in beta)
plot_modular_matrix(infomap_object)
```

### Infomap
Under the hood, the function `run_infomap_monolayer` runs:

```C++
./Infomap infomap.txt . --tree -i link-list --seed 123 -N 20 -f undirected -2
```

With this command, Infomap detects modules that contain both plants and pollinators.
* `-f undirected` indicates flow on an undirected network.
* `-2` indicates a two-level solution, with no hierarchical modules.


### Output
A [tree file](https://www.mapequation.org/infomap/#OutputTree) is produced by Infomap, but is parsed by `run_infomap_monolayer` from infomapecology (in R: `?run_infomap_monolayer`).
