---
title: "Monolayer directed network with hierarchical structure"
output:
  html_document: 
    fig_height: 4
    fig_width: 6
    theme: flatly
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      number_sections: true
---

```{r klippy, echo=FALSE, include=TRUE}
# Use this to add a copy-code buttong
klippy::klippy(position = c('top', 'right'), color = 'darkred')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data set
A binary directed marine food web from [Jacob, Ute, Aaron Thierry, Ulrich Brose, Wolf E. Arntz, Sofia Berg, Thomas Brey, Ingo Fetzer, et al. 2011. “The Role of Body Size in Complex Food Webs: A Cold Case.” In Advances in Ecological Research, edited by Andrea Belgrano, 45:181–223. Academic Press.](https://www.sciencedirect.com/science/article/pii/B9780123864758000058) Kongsfjorden is a glacial fjord on the northwest corner of the Svalbard archipelago. It is a 30 km open fjord with no marked sill at the entrance, and with a maximum depth exceeding 300 m. The network consists of 262 species with 1,544 feeding interactions.
Data is available [here](https://github.com/Ecological-Complexity-Lab/infomap_ecology_package/tree/master/code_from_paper)

### Input
A [link-list](https://www.mapequation.org/infomap/#InputLinkList).

### R Code
```{r monolayer_heirarchical_1}
# Import data
data("kongsfjorden_links")
data("kongsfjorden_nodes")
nodes <- kongsfjorden_nodes %>%
  select(node_name=Species, node_id_original=NodeID, everything())

interactions<- kongsfjorden_links %>%
  select(from=consumer, to=resource) %>%
  mutate_if(is.factor, as.character) %>%
  mutate(weight=1)

# Prepare network objects
network_object <- create_monolayer_network(x=interactions, directed = T, bipartite = F, node_metadata = nodes)

# Run infomap, allow hierarchical modules
# Some species will have only incoming or outgoing links, so the next line will result in a warning
infomap_object <- run_infomap_monolayer(network_object, infomap_executable='Infomap',
                                        flow_model = 'directed',
                                        silent=T,trials=100, two_level=F, seed=123)
```

### Infomap
Under the hood, the function `run_infomap_monolayer` runs:
```C++
./Infomap infomap.txt . -i link-list --tree --seed 123 -N 100 -f directed --silent
```
* Here, the most important feature is **the lack of the argument** `-2` (or `--two-level`), which allows for hierarchical modules (the default in Infomap).
* `-f directed` indicates flow on a directed network.

### Output
A [tree file](https://www.mapequation.org/infomap/#OutputTree) is produced by Infomap, but is parsed by `run_infomap_monolayer` from infomapecology. In Infomap's tree output, the `path` column has a tree-like format such as 1:3:2, which describes the path from the root of the tree to the leaf node. The first integer (1 in the 1:3:2 example) is the top module. The last integer after the colon (2 in the 1:3:2 example) indicates the ID of the leaf in the module, and not the ID of the node. In case a node has fewer levels because its module is not partitioned, the missing levels will show `NA`.

In the output excerpt below for the analysis of the Kongsfjorden food web, there are four nodes that belong to the 3rd submodule within top module 1. For these, `module_level3` is the leaf id. Top modules 2 and 3 do not have sub-modules. Therefore, for these modules the second level (`module_level2`) is the leaf id and `module_level3` shows an `NA` value. In addition, this data has node attributes, which are added to the final output. In this example: FeedingType, Mobility and Environment. 

<table style="border-collapse: collapse; width: 100%;">
  <tr>
    <th style="border: 1px solid #000; padding: 8px; text-align: left;">node_id</th>
    <th style="border: 1px solid #000; padding: 8px; text-align: left;">node_name</th>
    <th style="border: 1px solid #000; padding: 8px; text-align: left;">module_level1</th>
    <th style="border: 1px solid #000; padding: 8px; text-align: left;">module_level2</th>
    <th style="border: 1px solid #000; padding: 8px; text-align: left;">module_level3</th>
    <th style="border: 1px solid #000; padding: 8px; text-align: left;">FeedingType</th>
    <th style="border: 1px solid #000; padding: 8px; text-align: left;">Mobility</th>
    <th style="border: 1px solid #000; padding: 8px; text-align: left;">Environment</th>
  </tr>
  <tr>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">54</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">Laminaria saccharina</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">none</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">benthic</td>
  </tr>
  <tr>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">55</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">Laminaria digitata</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">2</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">none</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">benthic</td>
  </tr>
  <tr>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">57</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">Laminaria solidungula</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">3</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">none</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">benthic</td>
  </tr>
  <tr>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">49</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">Fucus distichus</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">4</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">none</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">benthic</td>
  </tr>
  <tr>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">22</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">Calanus finnmarchicus</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">2</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">predator</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">4</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">pelagic</td>
  </tr>
    <tr>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">28</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">Metridia lucens</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">1</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">2</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">2</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">predator</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">4</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">pelagic</td>
  </tr>
  <tr>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">102</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">Pandalus borealis</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">2</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">2</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">NA</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">predator</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">3</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">benthic</td>
  </tr>
  <tr>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">40</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">Thysanoessa inermis</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">2</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">3</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">NA</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">grazer</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">4</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">epipelagic/ice associated</td>
  </tr>
  <tr>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">26</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">Copepoda nauplii</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">3</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">2</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">NA</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">predator</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">4</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">pelagic</td>
  </tr>
  <tr>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">34</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">Clione limacina</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">3</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">3</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">NA</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">predator</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">4</td>
    <td style="border: 1px solid #000; padding: 8px; text-align: left;">pelagic</td>
  </tr>
</table>

    




## A second example with different data 

### Data set
A binary directed food web from [Mouritsen KN, Poulin R, McLaughlin JP, Thieltges DW. Food web including metazoan parasites for an intertidal ecosystem in New Zealand: Ecological Archives E092-173. Ecology. 2011;92: 2006–2006.](https://esajournals.onlinelibrary.wiley.com/doi/abs/10.1890/11-0371.1)

In infomapecology:
```R
data(otago_nodes)
data(otago_links)
```

### Input
A [link-list](https://www.mapequation.org/code.html#Link-list-format).

### R Code
```{r monolayer_heirarchical_2}
# Prepare data
otago_nodes_2 <- otago_nodes %>%
  filter(StageID==1) %>%
  select(node_name=WorkingName, node_id_original=NodeID, WorkingName,StageID, everything())
anyDuplicated(otago_nodes_2$node_name)
otago_links_2 <- otago_links %>%
  filter(LinkType=='Predation') %>% # Only include predation links
  filter(ConsumerSpecies.StageID==1) %>%
  filter(ResourceSpecies.StageID==1) %>%
  select(from=ResourceNodeID, to=ConsumerNodeID) %>%
  left_join(otago_nodes_2, by=c('from'='node_id_original')) %>%
  select(from, node_name, to) %>%
  left_join(otago_nodes_2, by=c('to'='node_id_original')) %>%
  select(from=node_name.x, to=node_name.y) %>%
  mutate(weight=1)

# Prepare network objects
# Some species will have only incoming or outgoing links, so the next line will result in a warning
network_object <- create_monolayer_network(x=otago_links_2, directed = T, bipartite = F, node_metadata = otago_nodes_2)

# Run infomap without hieararchy
infomap_object <- run_infomap_monolayer(network_object, infomap_executable='Infomap',
                                        flow_model = 'directed',
                                        silent=T,trials=100, two_level=T, seed=123)

infomap_object$modules %>%
  select(node_id, node_name, module=module_level1, OrganismalGroup, NodeType) %>%
  group_by(module, OrganismalGroup) %>% summarise(n=n_distinct(node_id)) %>% drop_na() %>%
  ggplot(aes(x=module, y=OrganismalGroup, size=n))+geom_point()+
  scale_x_continuous(breaks = 1:infomap_object$m)+
  theme_bw()+theme(panel.grid.minor = element_blank())


# Run infomap with hieararchy
infomap_object <- run_infomap_monolayer(network_object, infomap_executable='Infomap',
                                        flow_model = 'directed',
                                        silent=T,trials=100, two_level=F, seed=123)
infomap_object$modules %>%
  select(node_id, node_name, module=module_level1, OrganismalGroup, NodeType) %>%
  group_by(module, OrganismalGroup) %>% summarise(n=n_distinct(node_id)) %>% drop_na() %>%
  ggplot(aes(x=module, y=OrganismalGroup, size=n))+geom_point()+
  scale_x_continuous(breaks = 1:infomap_object$m)+
  theme_bw()+theme(panel.grid.minor = element_blank())
```

 
