---
title: "Temporal multilayer network with interlayer edges"
output:
  html_document: 
    fig_height: 4
    fig_width: 6
    theme: flatly
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      number_sections: true
---
```{r klippy, echo=FALSE, include=TRUE}
# Use this to add a copy-code buttong
klippy::klippy(position = c('top', 'right'), color = 'darkred')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data set
A temporal multilayer network. Each layer is a host-parasite bipartite network. Intralayer edges between a parasite species and a host species are the number of parasite individuals divided by the number of host individuals. Interlayer coupling edges connect each physical node to itself in the next layer (e.g., host A in layer 1 to host A in layer 2), and are calculated as the number of individuals in layer l+1 divided by the number of individuals in layer l. They therefore represent population dynamics. Interlayer edges only go one way (_l-->l+1_) because time flow one way. We represented the undirected edges within each layer as directed edges that go both ways (with the same weight) to be able to have a directed flow. This does not affect the calculation of L. This data set was taken from [_Pilosof S, Porter MA, Pascual M, Kéfi S. The multilayer nature of ecological networks. Nature Ecology & Evolution. 2017;1: 0101_](https://www.nature.com/articles/s41559-017-0101?proof=true19).

**Data sets in infomapecology:**

The data consists of several matrices representing the host-parasite bipartite network each year, an interlayer extended edge list defining the interlayer links between the same species in different years, and a table containing the attributes of the physical nodes (species). 

```{r multilayer_inter_1}
# Get data
data(siberia1982_matrix)
data(siberia1983_matrix)
data(siberia1984_matrix)
data(siberia1985_matrix)
data(siberia1986_matrix)
data(siberia1987_matrix)
data(siberia_interlayer)
data(siberia_nodes)
```

### Input
A [general multilayer link-list](https://www.mapequation.org/infomap/#InputMultilayer). This multilayer format gives full control of the dynamics, and no other movements are encoded. **When using this input format, it is expected that interlayer edges will be provided, otherwise there will be no inter-layer links in the final state network.**

```Python
*Multilayer
#layer node layer node weight
1 4 1 5 0.1585
1 4 1 6 0.2143
1 4 1 7 0.8276
... some more links...
4 20 4 74 0.0019
4 23 4 10 24.6
4 24 4 5 0.1143
... some more links...
2 16 3 16 1.14285714285714
4 29 5 29 3.66666666666667
5 57 6 57 0.841212121212121

```

### R Code
The description of functions `create_multilayer_network` and `run_infomap_multilayer` in the `emln` and `infomapecology` packages contains everything you need to know. Plotting is done using two dedicated functions: `plot_multilayer_modules` and `plot_multilayer_alluvial`.

```{r multilayer_inter_2}
# Create a multilayer object
layer_attrib <- tibble(layer_id=1:6, layer_name=c('1982','1983','1984','1985','1986','1987'))                                     

multilayer_siberia <- create_multilayer_network(list_of_layers = list(siberia1982_matrix,
                                                                    siberia1983_matrix,
                                                                    siberia1984_matrix,
                                                                    siberia1985_matrix,
                                                                    siberia1986_matrix,
                                                                    siberia1987_matrix),
                                              layer_attributes = layer_attrib,
                                              interlayer_links = siberia_interlayer,       
                                              bipartite = T,
                                              directed = F,physical_node_attributes = siberia_nodes )

#Run infomap
multilayer_siberia_modules <- run_infomap_multilayer(M=multilayer_siberia, relax = F, flow_model = 'directed', silent = T, trials = 100, seed = 497294, temporal_network = T)


#Module persistence
modules_persistence <- multilayer_siberia_modules$modules %>%
  group_by(module) %>%
  summarise(b=min(layer_id), d=max(layer_id), persistence=d-b+1) %>%
  count(persistence) %>%
  mutate(percent=(n/max(multilayer_siberia_modules$module$module))*100)

# 1. Modules' persistence
plot_multilayer_modules(multilayer_siberia_modules, type = 'rectangle', color_modules = T)+
  scale_x_continuous(breaks=seq(0,6,1))+
  scale_y_continuous(breaks=seq(0,70,5))+
  scale_fill_viridis_c()+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title = element_text(size=20),
        axis.text = element_text(size = 20),
        legend.text =  element_text(size=15),
        legend.title = element_text(size=20))

#2. Species flow through modules in time
plot_multilayer_alluvial(multilayer_siberia_modules, module_labels = F)+
  scale_x_continuous(breaks=seq(0,6,1))+
  scale_y_continuous(breaks=seq(0,70,5))+
  labs(y='Number of species')+
  theme_bw()+
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.text = element_text(color='black', size = 20),
        axis.title = element_text(size=20))
```

### Infomap
Under the hood, the function `run_infomap_multilayer` runs:
```C++
./Infomap infomap_multilayer.txt . -2 --seed 497294 -N 100 -i multilayer -f directed --silent
```

Explanation of key arguments:
* `-i multilayer` indicates a multilayer input format, which is automatically recognized as a [general multilayer link-list](https://www.mapequation.org/infomap/#InputMultilayer).
* `-f directed` indicates flow on a directed network. The visitation rates of nodes is obtained with a PageRank algorithm based on the direction and weight of edges. This includes interlayer edges.

### Output
For multilayer network the output file has a `_states` suffix, with the following format. Note the state_id column. For example, node 6 in layer 5 has a state_id of 377 (last line). The state_ids are created by Infomap but not used in our R code.

```python
# path flow name state_id node_id layer_id
1:1 0.0350053 "7" 334 7 6
1:2 0.0294496 "7" 205 7 4
1:3 0.0291166 "7" 280 7 5
1:4 0.0283014 "7" 137 7 3
1:5 0.0124629 "7" 69 7 2
1:6 0.00734334 "7" 3 7 1
1:7 0.00180819 "33" 171 33 3
         ...
11:1 0.00929888 "52" 328 52 6
11:2 0.00861513 "52" 272 52 5
11:3 0.00694999 "52" 211 52 4
11:4 0.00564292 "52" 146 52 3
11:5 0.00256943 "52" 76 52 2
11:6 0.00101397 "52" 14 52 1
12:1 0.0083305 "16" 340 16 6
         ...
```

This output is parsed by `run_infomap_multilayer` to obtain a table in which each state node (combination of a physical node in a layer) is assigned to a module. This can be obtained by:
```{r multilayer_inter_3}
multilayer_siberia_modules$modules %>% select(node_id, layer_id, node_name, type, module)
```
