---
title: "Temporal multilayer network with global relax rates"
output:
  html_document: 
    fig_height: 4
    fig_width: 6
    theme: flatly
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      number_sections: true
---
```{r klippy, echo=FALSE, include=TRUE}
# Use this to add a copy-code buttong
klippy::klippy(position = c('top', 'right'), color = 'darkred')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data set
Load the 65th multilayer network from the [_`emln package`_](https://github.com/Ecological-Complexity-Lab/emln). This particular network is a temporal plant pollinator multilayer network without interlayer interactions, and it consists of 6 layers, corresponding to the flowering season between 2006 and 2011.

```{r multilayer_emln_1}
# Load multilayer network number 65
emln_65 <- load_emln(65)
emln_65$layers
head(emln_65$nodes)
head(emln_65$extended)
```
### Input
A [multilayer link-list format](https://www.mapequation.org/infomap/#InputMultilayer). With this format, a random walker moves within a layer and with a given relax rate _r_ jumps to another layer without recording this movement, such that the constraints from moving in different layers can be gradually relaxed. If the `*Inter` exists, then interlayer edges can be used. However, if a relax rate is also used (Infomap input parameter `--multilayer-relax-rate`), the interlayer edges are ignored. Here we only include `*Intra`.

```Python
# A network in multilayer format
*Intra
#layer node node weight
1 101 1 19
1 113 1 2
1 152 1 1
... some more cases ...
6 69 11 17
6 70 11 2
6 87 11 3
```

### R Code
The description of functions `create_multilayer_object` and `run_infomap_multilayer` in the `infomapecology` package contains everything you need to know. The `for` loop performs a sensitivity analysis to examine how structure changes with increasing relax rates.

```{r multilayer_emln_2}
#Run Infomap and return the network's modular structure at increasing relax-rates.
relaxrate_modules <- NULL
for (r in seq(0.001,1, by = 0.0999)){
  print(r)
  modules_relax_rate <- run_infomap_multilayer(emln_65, relax = T, silent = T, trials = 50, seed = 497294, multilayer_relax_rate = r, multilayer_relax_limit_up = 1, multilayer_relax_limit_down = 0, temporal_network = T)
  modules_relax_rate$modules$relax_rate <- r
  relaxrate_modules <- rbind(relaxrate_modules, modules_relax_rate$modules)
}
```

### Infomap
Under the hood, the function `run_infomap_multilayer` runs:
```C++
./Infomap infomap_multilayer.txt . -2 --seed 497294 -N 50 -i multilayer --multilayer-relax-rate 0.15 --multilayer-relax-limit-up 1 --multilayer-relax-limit-down 0 --silent
```

Explanation of key arguments:
* `-i multilayer` indicates a multilayer input format, which is automatically recognized as a multilayer link list (and not as a general link list).
* `--multilayer-relax-rate 0.15` defines the relax rate (here 0.15).
* `multilayer-relax-limit-up 1 --multilayer-relax-limit-down 0` limits relaxation to a single layer upwards (l to l+1), but never downwards, because time flows one-way.

### Output
For multilayer network the output file has a `_states` suffix, with the following format. This output is as in [Temporal multilayer network with interlayer edges](multilayer_interlayer.md). 

```Python
# path flow name stateId physicalId layerId
# path flow name state_id node_id layer_id
1:1 0.0208474 "12" 604 12 6
1:2 0.0166813 "100" 531 100 6
1:3 0.0103777 "45" 600 45 6
1:4 0.00927889 "58" 526 58 6
```
This output is parsed by `run_infomap_multilayer` to obtain a table in which each state node (combination of a physical node in a layer) is assigned to a module. This can be obtained by (after running the code above):
```{r multilayer_emln_3}
#relaxrate_modules$modules
relaxrate_modules %>% select(node_id, layer_id, node_name, type, module)
```
